---
title: Maketing Campaign Analysis
author: Audrey Nguyen, Shado Yang, Cecile Wiederkehr, Chengliang Fu, Edward Zhang, Weibin Huo
date: April 2022
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE,warning = FALSE)
```

## Introduction

The Goal of this project is to analyze the Customer Personality data to help businesses better understand their customers and allows them to develop targeted marketing campaigns according to the specific needs, behaviors and concerns of different types of customers. This analysis can identify which customer segment is most likely to buy a product and then market the product only on that particular segment.


```{r,results='hide',message=FALSE}
#load libraries
library(tidyverse)
library(broom)
library(kableExtra)
library(psych)
library(pixiedust)
library(caret)
library(fastDummies)
library(factoextra)
library(cluster)
library(ggplot2)
```


```{r, include=FALSE}
#Load up dataset 

library(readr)
mktg <- read_delim("marketing_campaign.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

```


```{r, include=FALSE}

# Remove NA
mktg1 = na.omit(mktg)
```
## Data

The data consists of the following customer attributes:

* Customer's birth year from which the age can be derived

* Customer's education level that has been divided into 2 categories: graduates and non-graduates

* Customer's marital status that has been divided into 2 categories: single and non-single

* Customer's yearly household income

* Number of children in customer's household

* Number of teenagers in customer's household

* Date of customer's enrollment with the company

* Number of days since customer's last purchase

* Complain: 1 if the customer complained in the last 2 years, 0 otherwise

The data also gives information on how much each customer has spent on wine, fruits, meat products, fish products, sweet products and gold products in the last 2 years.

Their is also an indication wheather the customers have responded to 5 different campaigns.

```{r, include=FALSE}
#Create new columns for Customer Age, No. of Child in House, Total Spending and Whether they Accepted Marketing efforts or not

mktg1['Age']=2022-mktg1$Year_Birth
mktg1['Children']=mktg1$Kidhome+mktg1$Teenhome
mktg1['Accepted']=mktg1$AcceptedCmp1+mktg1$AcceptedCmp2+mktg1$AcceptedCmp3+mktg1$AcceptedCmp4+mktg1$AcceptedCmp5
mktg1['TotalSpent']=mktg1$MntMeatProducts+mktg1$MntFishProducts+mktg1$MntWines+mktg1$MntFruits+mktg1$MntSweetProducts+mktg1$MntGoldProds

head(mktg1)
mktg2<-mktg1
```

```{r, include=FALSE}
max(mktg1$Income) #666'666
min(mktg1$Income) #1'730
mean(mktg1$Income) #52'247
nrow(mktg1[mktg1$Income>125000,]) #8
```
The income ranges between $1'730 and $666'666. Only 8 customers have incomes over $125,000, the rest are normaly distributed.

They're are customers with ages greater than 120 years old, which could be a mistake so we did not include them in the analysis.

```{r, include=FALSE}
mktg1 <- mktg1[mktg1$Age<=100,]
mktg1 <- mktg1[mktg1$Income<125000,]
```

```{r, echo=FALSE, width=2.5}
#Histogram of Income
hist(mktg1$Income,40, xlab = "Income", main= "Histogram of Income",xlim=c(0,125000))
```


```{r, include=FALSE}
hist(mktg1$Age,40, xlab = "Age", main= "Histogram of Age")

```

```{r, include=FALSE}
mktg1fortable = mktg1 %>% rename(First=AcceptedCmp1,
                              Second=AcceptedCmp2,
                              Third=AcceptedCmp3,
                              Fourth=AcceptedCmp4,
                              Fifth=AcceptedCmp5)
mktg1fortable <- mktg1fortable[, c(24, 25, 21, 22, 23)]

sumcamp = mktg1fortable %>% summarize(first = sum(First),second = sum(Second), third = sum(Third), fourth = sum(Fourth), fifth = sum(Fifth))# first 142, second 30, third 163, fourth 164, fifth 162
table <- round(cor(mktg1fortable),2)
```


```{r, include=FALSE}
t.test(mktg1fortable$First, mktg1fortable$Second,alternative= "greater") #p-value=0
```

## Analysis

### Campaign responses

Overall, the most popular campaigns were the third fourth and fifth. The second campaign performed fatalistically significantly worse than all other campaigns.

```{r, echo=FALSE}
knitr::kable(sumcamp,  "simple", caption = "Number of Reponses to each Campaign") 
```


Some customers responded to several campaigns. The respondents to the first campaign correlate to the ones in the fifth campaign. There is hardly any correlation between the third campaign and the second and the third and the fourth.


```{r, echo=FALSE}
knitr::kable(table,  "simple", caption = "Correlation between different Campaigns") 
```


### Prediction

High amount spent on Wine, Gold and Meat Products make them more likely to respond.


```{r, include=FALSE}
#covert marital status
mktg1 <- mktg1 %>% mutate(Marital_Status = replace(Marital_Status, Marital_Status == "Divorced" | Marital_Status == "Widow" | Marital_Status == "Alone" | Marital_Status == "Absurd" | Marital_Status == "YOLO", "Single"))
mktg1 <- mktg1 %>% mutate(Marital_Status = replace(Marital_Status, Marital_Status == "Together" | Marital_Status == "Married", "Taken"))

mktg1 <- mktg1 %>% mutate(Education = replace(Education, Education == "Graduation"| Education == "PhD" | Education == "Master", "graduate"))
mktg1 <- mktg1 %>% mutate(Education = replace(Education, Education == "Basic"| Education == "2n Cycle", "non-graduate"))
str(mktg1)
```


```{r, include=FALSE}
out2 <- glm(mktg1$Response ~ MntWines + MntFruits + MntMeatProducts + MntFishProducts + MntSweetProducts + MntGoldProds, data = mktg1, family = "binomial")
summary(out2)
#Wine, Gold and Meat Products make them more likely to respond

out3 <- glm(mktg1$MntWines ~ Education + Marital_Status + Income + Age + Kidhome + Teenhome, data = mktg1, family = "gaussian")
summary(out3)

out4 <- glm(mktg1$MntGoldProds ~ Income + Age + Kidhome + Teenhome, data = mktg1, family = "gaussian")
summary(out4)
```

### Clustering

###PCA
```{r}
library(corrplot)
library(collapse)
library(bestNormalize)
library(recipes)
library(dplyr)
library(reshape2)
library(ggplot2)

df6<- mktg2[,-c(1,2,6:8,27,28)]
df6$Education<- as.factor(df6$Education)
#skim(df6)
###################################################################
#Trasform Categorical to numeric

#Education
df6$Education <- fct_collapse(df6$Education,
                                         "3" = c("2n Cycle", "Master","PhD"),
                                         "1" = "Basic",
                                         "2" = "Graduation")
df6$Education <- as.numeric(levels(df6$Education))[df6$Education]

#Marital_Status
#combine Absurd,Alone,Divorced,Single,Widow and YOLO into one category of Single,and Married and Together into Together. Then I will numerically encode these categories as 1 and 2 respectively.
df6$Marital_Status <- fct_collapse(df6$Marital_Status,
                                         "1" = c("Absurd", "Alone","Divorced","Single",
                                               "Widow","YOLO"),
                                         "2" = c("Married","Together"))
df6$Marital_Status <- as.numeric(levels(df6$Marital_Status))[df6$Marital_Status]

df6 %>% 
  cor() %>% 
  corrplot( tl.cex = .7, tl.col = "black", method = "ellipse")
```
```{r}
pca <- recipe(~., data = df6)
#PCA is performed after standardizing each numeric variable to have mean zero and standard deviation one. If we dont do that then the variable with high value of mean has higher influence in the principal components generated.
pca1 <- pca %>% step_orderNorm(all_numeric_predictors())
pca <- pca1 %>% step_normalize(all_numeric_predictors())
summary(pca)
# apply principal component analysis on the dataset processed so far and see how many principal components explain most of the variance.
check_pca <- pca %>% 
  prep() %>% 
  bake(new_data = NULL) %>% 
  prcomp()
check_pca

tidy(check_pca, matrix = "eigenvalues")
#Retain enough to explain a "large enough" proportion of the original data, 70% to 90% is usually used, so the suggest component is 8 

pca <- pca %>% 
  step_pca(all_numeric_predictors(), num_comp = 8) %>% 
  prep() %>%
  bake(new_data=NULL)
```

```{r}
##Cluster Analysis for pca
library(FactoMineR)
library(factoextra)
#Creating a factor map for the top 8 variables with the highest contributions
staging <- PCA(df6, graph = FALSE)
fviz_pca_var(staging, col.var = "contrib", gradient.cols = c("#002bbb", "#bb2e00"),select.var = list(contrib = 8), repel = TRUE)

fviz_contrib(staging, choice = "var", axes = 1, top = 8)
fviz_pca_biplot(staging)

fviz_pca_var(staging, col.var = "contrib", gradient.cols = c("#002bbb", "#bb2e00"), repel = TRUE)
```




```{r, include=FALSE}
mktg1=mktg1[c(-1,-2,-6,-7,-8,-20,-21,-22,-23,-24,-25)]
```

```{r, include=FALSE}
view(mktg1)
```


```{r, include=FALSE}
head(mktg1)
```

```{r, include=FALSE}
#create dummy for education and marital_status

dmy <- dummyVars(" ~ .", data = mktg1, fullRank = T)
dat_transformed <- data.frame(predict(dmy, newdata = mktg1))

glimpse(dat_transformed)
```

```{r, include=FALSE}
df1=dat_transformed[c(-16,-17)]
```

```{r, include=FALSE}
str(df1)
```


```{r, include=FALSE}
set.seed(123)
df1.scaled = scale(df1)
kfit <- kmeans(df1.scaled, center=3)
```

```{r, include=FALSE}
#total variance
kfit$tot.withinss
```

```{r, include=FALSE}
#size of each cluster
kfit$size

```

```{r, include=FALSE}
screedf = data.frame(k = 1:10,tot.withinss = NA)
for(k in 1:nrow(screedf)){
screedf$tot.withinss[k] =
kmeans(df1.scaled,centers = k)$tot.withinss
}
ggOut = ggplot(screedf,aes(k,tot.withinss)) +
geom_line() + theme_bw(15) + geom_point() +
scale_x_continuous(breaks = 1:10) +
xlab("Number of Clusters") +
ylab("Total Within Group Sum of Squares")

```

```{r, include=FALSE}
ggOut
```

```{r, include=FALSE}
x<-fviz_nbclust(df1.scaled,kmeans,method="wss")+geom_vline(xintercept=3,linetype=2)
x
```


```{r, include=FALSE}
print(kfit$betweenss/kfit$totss)
```

```{r, include=FALSE}
fviz_cluster(kfit, df1, geom = "point",ellipse.type = "norm",repel = TRUE)
```

```{r, include=FALSE}
clusplot(df1.scaled,kfit$cluster,
color = TRUE,
main = "k=3")

```


```{r, include=FALSE}
library(dbplyr)
df_new<-cbind(df1, kfit$cluster)
df_new$`kfit$cluster`<-as.factor(df_new$`kfit$cluster`)

```

```{r, include=FALSE}
df1['cluster']=as.factor(kfit$cluster)
```

```{r, include=FALSE}
head(df1)
```

```{r, echo=FALSE}
view(df1)
```


```{r, echo=FALSE}
#identify cluster
ggplot(df1, aes(x=TotalSpent, y=Income, color=cluster)) +
  geom_point()
```

```{r}
acceptplot = ggplot(df1, aes(y = Accepted)) +
 geom_bar(aes(fill = cluster), position = position_stack(reverse = TRUE)) +
 theme(legend.position = "top")
acceptplot
```



```{r, include=FALSE}
wineplot = ggplot(df1, aes(x=cluster,y=MntWines,fill=cluster))+geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=2, notch=T)
wineplot
```

```{r, include=FALSE}
fruitplot = ggplot(df1, aes(x=cluster,y=MntFruits,fill=cluster))+geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=2, notch=T)
fruitplot
```

```{r, include=FALSE}
meatplot = ggplot(df1, aes(x=cluster,y=MntMeatProducts,fill=cluster))+geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=2, notch=T)
meatplot
```

```{r, include=FALSE}
fishplot = ggplot(df1, aes(x=cluster,y=MntFishProducts,fill=cluster))+geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=2, notch=T)
fishplot
```

```{r, include=FALSE}
sweetplot = ggplot(df1, aes(x=cluster,y=MntSweetProducts,fill=cluster))+geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=2, notch=T)
sweetplot
```

#Extra
```{r}
wss <- function(k) {
  kmeans(pca , k, nstart = 200 )$tot.withinss
}
fviz_nbclust(pca, kmeans, method = "wss")

res_kmeans <- kmeans(pca, centers = 3, nstart = 200)
marketing_clustered <- df6 %>% mutate(cluster = res_kmeans$cluster)

```

#Eduaction
```{r}
explore <- marketing_clustered
explore$Education <- as.factor(explore$Education)
explore$cluster <- as.factor(explore$cluster)
ggplot(data = explore) + geom_jitter(mapping = aes(x=cluster,y=Education))

#Only cluster 3 has individuals with basic education
```

#Marital_Status
```{r}
explore <- marketing_clustered
explore$Marital_Status <- as.factor(explore$Marital_Status)
explore$cluster <- as.factor(explore$cluster)
ggplot(data = explore) + geom_jitter(mapping = aes(x=cluster,y=Marital_Status))

#we can tell nothing 
```
#income
```{r}


ggplot(explore) + geom_histogram(mapping = aes(x=Income),bins=50) + facet_wrap(~cluster)
```

#platform
```{r}

marketing_clustered%>%
  select(cluster, NumWebPurchases, NumCatalogPurchases, NumStorePurchases, NumDealsPurchases, NumWebVisitsMonth)%>%
  melt(id='cluster')%>%
  ggplot(aes(as_factor(cluster), value))+
  geom_boxplot()+
  facet_wrap(~variable, ncol = 5)
```


#Amount spent on different items
```{r}
marketing_clustered%>%
  select(cluster, MntWines, MntFruits, MntMeatProducts, MntSweetProducts, MntGoldProds)%>%
  melt(id='cluster')%>%
  ggplot(aes(as_factor(cluster), value))+
  geom_boxplot()+
  facet_wrap(~variable, ncol=5)

#Most of the amount is spent on wines and meats. Cluster 1 and 3 are high spenders which is directly correlated with income
```

##AD
```{r}
marketing_clustered %>% 
                  select(AcceptedCmp1,AcceptedCmp2,AcceptedCmp3,AcceptedCmp4,AcceptedCmp5,Response,cluster) %>%
                  group_by(cluster) %>%
                  summarise(AcceptedCmp1 = sum(AcceptedCmp1),
                            AcceptedCmp2 = sum(AcceptedCmp2),
                            AcceptedCmp3 = sum(AcceptedCmp3),
                            AcceptedCmp4 = sum(AcceptedCmp4),
                            AcceptedCmp5 = sum(AcceptedCmp5),
                            Response = sum(Response))
```

